
The experimental discrepancy with the Standard Model prediction for
the muon anomalous magnetic moment heightens interest in the
possibility of measuring $g-2$ of the $\tau$ lepton using angular
distributions in $\tau$-pair production. This can be done at a super flavor factory, with or without electron polarization.  

The best current bound on the $\tau$ anomalous moment  $a_{\tau}=(g-2)/2$ is indirect, derived from the LEP2 measurement of the total cross section for $e^+e^- \rightarrow e^+e^- \tau^+\tau^-$: $-0.052 < a_\tau < 0.013$ @ 95\% CL. This is well above the Standard Model prediction: $a_\tau^{SM} = 1177.21(5) \times 10^{-6}$, but even so provides a model-independent bound on New Physics contributions:  $-0.007 < a_\tau^{NP} < 0.005$ @ 95\% CL.~\cite{ref:bern2}

This measurement can be done in $e^+ e^- \rightarrow \tau^+\tau^-$ production with unpolarized beams. Determination of the real part of the form factor requires the measurement of correlations between the decay products of both polarized taus. 

With polarized taus one can access new observables, $A_T^\pm$ and $A_L^\pm$, the transverse and longitudinal polarizations of the outgoing $\tau$s,
\[\begin{array}{l}
A_T^ \pm \, = \,\frac{{\sigma _R^ \pm |{P_e}\,\, - \,\sigma _L^ \pm |{P_e}}}{\sigma }\, = \, \mp \,{\alpha _ \pm }\frac{{3\pi }}{{8(3 - {\beta ^2})\gamma }}\left[ {{{\left| {{F_1}} \right|}^2}\, + \,(2 - {\beta ^2}){\gamma ^2}{\mathop{\rm Re}\nolimits} \left\{ {{F_2}} \right\}} \right]\\
\,\,A_L^ \pm \, = \,\frac{{\sigma _{{\rm{FB}}}^ \pm ( + )|{P_e}\, - \,\sigma _{{\rm{FB}}}^ \pm ( - )|{P_e}}}{\sigma }\, = \, \mp {\alpha _ \pm }\frac{3}{{4(3 - {\beta ^2})}}\left[ {{{\left| {{F_1}} \right|}^2}\, + \,2{\mathop{\rm Re}\nolimits} \{ {F_2}\} } \right]\\
{\rm{                  Re}}\left\{ {{F_2}(s)} \right\}\, = \, \mp \frac{{8(3 - {\beta ^2})}}{{3\pi \gamma {\beta ^2}}}\frac{1}{{{\alpha _ \pm }}}\left( {A_T^ \pm \, - \,\frac{\pi }{{2\gamma }}A_L^ \pm } \right).
\end{array}\]
that are estimated by
Bernab\'eu {\it et al.}\cite{ref:bern1} to increase the sensitivity to
$\Re(F_2)$ by a factor of three, to $\sim2\times 10^{-6}$ with 80\%
electron polarization, which could allow a measurement of the Standard
Model moment to a precision of several percent with a data sample of
75 ab$^{-1}$. 

Observation of a $\tau$ EDM would be evidence of $T$ violation.  $T$-odd observables can be isolated by the study of $\tau$ angular
distributions using unpolarized beams. The best current limit is from Belle~\cite{bellemom}:
\vskip -12pt
\[ - 0.22\,e{\rm{cm}}\,\, < \,{\mathop{\rm Re}\nolimits} \{ d_\tau ^\gamma \} \, \times \,{10^{16}}\, < \,0.45\,e{\rm{cm}}\,@\,95\% \,\,C\!L.\]
\vskip -8pt
Having a polarized electron
beam allows these investigations to be done using the decay products
of individual polarized taus. We can define a new, more sensitive $C\!P$-odd $T$-odd observable:
\vskip -12pt
\[A_N^{C\!P}\, = \,\frac{1}{2}\left( {A_N^ + \, + \,A_N^ - } \right)\, = \,{\alpha _h}\frac{{3\pi \gamma \beta }}{{8(3 - {\beta ^2})}}\frac{{2{m_\tau }}}{e}{\mathop{\rm Re}\nolimits} \left\{ {d_\tau ^\gamma } \right\}\]
where the azimuthal asymmetry for the two polarizations is 
\[A_N^ \mp \, = \,\frac{{\sigma _L^ \mp \, - \,\sigma _R^ \mp }}{\sigma }\, = {\alpha _ \mp }\frac{{3\pi \gamma \beta }}{{8(3 - {\beta ^2})}}\frac{{2{m_\tau }}}{e}{\mathop{\rm Re}\nolimits} \left\{ {d_\tau ^\gamma } \right\}\]
\vskip -12pt
The upper-limit sensitivity for the real part of the
$\tau$ EDM has been estimated to be to be $|\Re{d_\gamma}|\simeq 3\times 10^{-19} \ e \cdot {\rm cm}$
with 50 ab$^{-1 }$ at Belle II and $|\Re{d_\gamma}|\simeq 7\times 10^{-20} \ e \cdot {\rm cm}$
with 75 ab$^{-1 }$ with a polarized electron beam at Super$B$\cite{GonzalezSprinberg:2000mk}.

A $C\!P$-violating asymmetry in $\tau$ decay would be manifest evidence
for physics beyond the Standard Model. \babar\ has recently published
a 3$\sigma$ asymmetry in $\tau\to\pi K_S^0(\ge 0\pi^0)$
decay\cite{BABAR:2011aa}. The super flavor factories have the sensitivity to
definitely confirm or refute this measurement, and, further, provide
access to new $C\!P$-odd observables that increase the sensitivity in the
search for a $C\!P$ asymmetry to the level of $\sim 10^{-3}$.
